\sysfull

### 2017-11-13 paper/intro.tex
fixing a fig
Pass over Alexei revision of intro

\begin{figure}
	%\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
    \caption{\label{fig:coalesce}Task decomposition is required for
intermittent execution but task transitions introduce overhead.}
\end{figure}

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow.
%, which may be conditionally input-dependent.
%
Figure~\ref{fig:decomp} illustrates how the decomposition into tasks affects
the time the application takes to complete.
%
Each task transition introduces the overhead of tracking and atomically
committing the modifications to non-volatile memory by the task, to maintain
consistency of program state.
%
Naively minimizing the number of task transitions, (i.e. the task count),
\emph{at compile time} risks creating a task that requires more energy than the
device can buffer in its storage capacitor.
%
Such a large task would fail to complete unless the environment provides
sufficient incoming energy to supplement the energy in the capacitor.
%
To eliminate this risk, the programmer (or the compiler) must decompose the
program conservatively into many small tasks, at the cost of higher task
transition overhead.

One approach to decrease the task count, without assumming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size.
%
Howerver, statically predicting energy of arbitrary input-dependent code with
peripheral access is a problem without a general solution.
%
Furthermore, a static decomposition approach cannot produce decompositions
portable across devices with different storage capacitorss.
%
In this paper, we propose a \emph{dynamic} task decomposition approach, that
accepts any static decomposition and opportunistically \emph{coalesces} tasks
at runtime when there is sufficient energy.

@@

\textbf{Task Coalescing.} To add support for dynamic task
dvision to a task-based programming model, we propose a
\emph{task coalescing} mechanism: merging of atomic tasks
into a single (larger) task while preserving atomicity.
Merging two atomic tasks might produce a non-atomic task,
that may leave memory inconsistent after a partial
execution. A merge breaks atomicity when the accesses to
non-volatile memory from the previously atomic merged tasks
form a \emph{write-after-read (WAR)} dependency.
%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
Although WAR dependencies introduced by the \emph{static}
task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
run-time cannot.
%
To maintain memory consistency in merged tasks, we propose a
mechanism for \emph{memory virtualization} that buffers the
updates to non-volatile variables in volatile memory before
committing them at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.
%

Task coallescing removes the optimization burden of choosing
the best task decomposition from the programmer, because it
accepts any decomposition and improves it dynamically.
%
To further reduce the programmer effort, we propose a
compiler pass for \emph{automatic decomposition} of programs
into (small) \emph{atomic} tasks.
%
The compiler identifies non-volatile variables shared across
tasks as tasks are created, and instruments reads and writes
of those variables, using memory virtualization, to keep
the data consistent in the presence of power loss.
%
Despite being limited to a subset of the C language, the
automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate
effort.

We develop \sys%
\footnote{To be released at \url{http://anonymized.link}}
%
: a task coallescing system for task-based
programming models on intermittently-powered devices. The
capabilities required for effective coallescing mark our
contributions:

\begin{itemize}
\item Two adaptive coallescing strategies for selecting boundaries to coallesce based on past execution behavior
\item Software memory virtualization for transiently-powered devices that preserves atomicity of coallesced tasks
\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coallescing at runtime
\end{itemize}

>>>>

\begin{figure}
	%\includegraphics[width=\columnwidth]{figures/}
    \caption{\label{fig:coalesce}Task decomposition is required for intermittent execution but task transitions introduce overhead.}
\end{figure}

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. Figure~\ref{fig:coalesce} illustrates how the decomposition into tasks affects
the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor. Such a large task would fail to complete unless the environment provides
sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach cannot produce decompositions portable across devices with different storage capacitors.

@@

\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition approach, that accepts any static decomposition and opportunistically \emph{coalesces} tasks at runtime when there is sufficient energy. The \emph{task coalescing} is a mechanism where atomic tasks are merged into a single (larger) task while preserving atomicity.

Unfortunately, merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution. A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
run-time cannot. To maintain memory consistency in merged tasks, we propose a
mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.

Task coalescing removes the optimization burden of choosing the best task decomposition from the programmer, because it accepts any decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices. The capabilities required for effective coalescing mark our contributions:

\begin{itemize}
	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coalescing at runtime.
\end{itemize}
